---
title: "zearn_nudge"
format: pdf
---

```{r import}
library(tidyverse)
library(zoo)
library(gt)
library(fastICA)
set.seed(794563797)
# Timestamp: 2023-12-21 21:42:25 UTC
# © 1998-2023 RANDOM.ORG

teacher_usage <- read.csv("Data/raw/Teacher Usage - Time Series 2020-10-09T1635.csv")
classroom_student_usage <- read.csv("Data/raw/Classroom Student Usage - Time Series 2020-10-09T1616.csv")
classroom_info <- read.csv("Data/raw/Classroom Info 2020-10-09T1617.csv")
classroom_teacher_lookup <- read.csv("Data/raw/Classroom-Teacher Lookup 2020-10-09T1618.csv")
school_info <- read.csv("Data/raw/School Info 2020-10-09T1619.csv")
la_usage_types <- read_csv("Data/raw/la_usage_types.csv")
```

## Methods

### Study 1

#### Data Collection

  -   Source and timeframe of data.
  -   Specific teacher and student metrics collected.
  -   Data cleaning and preprocessing steps.
  -   Depositing of Raw Data: Ensure all raw data is available in a publicly accessible database.
  -   Criteria for selecting and excluding data points or subjects.

This dataset forms the basis for assessing the impact of our upcoming experimental intervention aimed at enhancing teaching methodologies and student learning outcomes.

```{r preprocessing}
teacher_usage <- teacher_usage %>%
  mutate(
    # Noting week, month, and year to summarize teacher behavior:
    week = lubridate::week(Usage.Time),
    year = lubridate::year(Usage.Time),
    wday = lubridate::wday(Usage.Time),
    hour = lubridate::hour(Usage.Time),
    whour = (wday - 1)*24 + hour,
    # Transforming Event.Type for "Resource Downloaded"
    Event.Type = ifelse(Event.Type == "Resource Downloaded",
                        paste0("RD.", Curriculum.Resource.Category),
                        Event.Type)
    )

# Summarize teacher behavior
teacher_usage_total <- teacher_usage %>%
  count(Adult.User.ID, Event.Type, week, year) %>%
  pivot_wider(names_from = Event.Type, values_from = n,
              values_fill = list(n = 0)) %>%
  full_join(teacher_usage %>%
              group_by(Adult.User.ID, week, year) %>%
              summarize(Minutes.on.Zearn...Total = sum(Minutes.on.Zearn...Total)),
            by = c("Adult.User.ID", "week", "year"))

# Remove teachers inactive for more than 2 months
active_teachers <- teacher_usage_total %>%
  count(Adult.User.ID) %>%
  filter(n > 8) %>%
  pull(Adult.User.ID)

# Merge teacher and student data
# Merging with the classroom info for further merge with teacher info
teacher_student_usage_subset <- classroom_student_usage %>%
  inner_join(classroom_teacher_lookup %>%
               # Filtering active teachers
               filter(Teacher.User.ID %in% active_teachers),
             by = "Classroom.ID") %>%
  mutate(week = lubridate::week(Usage.Week),
         year = lubridate::year(Usage.Week)) %>%
  inner_join(teacher_usage_total,
             by = c("Teacher.User.ID" = "Adult.User.ID",
                    "week", "year"))

# Adding the number of classrooms each teacher has
teacher_student_usage_subset <- teacher_student_usage_subset %>%
  inner_join(classroom_teacher_lookup %>%
              count(Teacher.User.ID) %>%
              rename(teacher.number.classes = n),
            by = "Teacher.User.ID")

# Adding the classrooms types
teacher_student_usage_subset <- teacher_student_usage_subset %>%
  inner_join(classroom_info, by = "Classroom.ID") %>%
  mutate(
    Grade.Level = factor(Grade.Level, ordered = TRUE)
    )

# Process la_usage_types
la_usage_types <- la_usage_types %>%
  mutate(
    curriculum = as.integer(`Schools and Districts Usage Type` %in%
                              c("20-21 Curriculum", "Core Complement"))
    )

# Further data merging
teacher_student_usage_subset <- teacher_student_usage_subset %>%
  left_join(la_usage_types %>%
              mutate(`Schools and Districts MDR School ID` =
                       as.integer(`Schools and Districts MDR School ID`)),
            by = c("MDR.School.ID" = "Schools and Districts MDR School ID")) %>%
  select(!`Schools and Districts Usage Type`) %>%
  left_join(school_info %>% 
              select(MDR.School.ID, Demographics...Zipcode.Median.Income),
            by = "MDR.School.ID") %>%
  mutate(Demographics...Zipcode.Median.Income =
           factor(Demographics...Zipcode.Median.Income,
                  ordered = TRUE))

# Processing students' data
teacher_student_usage_subset <- teacher_student_usage_subset %>%
  mutate(month = month(Usage.Week)) %>%
  # Remove June/July/August
  filter(!month %in% c(6, 7, 8))

# Filtering active classrooms
active_classrooms <- teacher_student_usage_subset %>%
  count(Classroom.ID) %>%
  # Remove classes inactive for more than 7 months
  filter(n > 20) %>%
  pull(Classroom.ID)

teacher_student_usage_subset <- teacher_student_usage_subset %>%
  filter(Classroom.ID %in% active_classrooms)

average_student <- teacher_student_usage_subset %>%
  # More than 5 students
  filter(Students...Total > 5) %>%
  group_by(Classroom.ID) %>%
  summarise(Active.Users...Total = mean(Active.Users...Total)) %>%
  # More than 5 active students on average
  filter(Active.Users...Total > 5)

teacher_student_usage_subset <- teacher_student_usage_subset %>%
  filter(Classroom.ID %in% average_student$Classroom.ID)

# Remove duplicate classroom-week pairs (classrooms with more than one teacher)
duplicated_classrooms <- teacher_student_usage_subset %>%
  mutate(unique_id = paste(Classroom.ID, Usage.Week)) %>%
  group_by(Classroom.ID) %>%
  # Identify Classroom.IDs with at least one duplicated Classroom-Week pair
  filter(any(duplicated(unique_id))) %>%
  distinct(Classroom.ID) %>%
  ungroup()

teacher_student_usage_subset <- teacher_student_usage_subset %>%
  # Filter out all rows with those Classroom.IDs
  filter(!Classroom.ID %in% duplicated_classrooms$Classroom.ID) %>%
  # Replace NAs with 0 and remove small classrooms
  mutate(Badges.per.Active.User = ifelse(is.na(Badges.per.Active.User),
                                         0, Badges.per.Active.User))
```
  
#### Summary Statistics

```{r}
#| label: tbl-summary
#| tbl-cap: "Pre-experimental Intervention Data. This table summarizes key educational metrics for Zearn teachers. The data was collected from July 2019 to June 2020."

# Pre-calculate unique teachers and classrooms
unique_teachers <- unique(teacher_student_usage_subset$Teacher.User.ID)
unique_classrooms <- unique(
  teacher_student_usage_subset[, c('Teacher.User.ID', 'Classroom.ID')]
  )

# Statistics for "# of classes"
n_teachers <- length(unique_teachers)
n_classes <- nrow(unique_classrooms)
mean_classes <- n_classes / n_teachers
classes_per_teacher <- unique_classrooms %>% count(Teacher.User.ID)

# Statistics for "# of badges"
teacher_student_usage_subset$Total.Badges <- 
  teacher_student_usage_subset$Active.Users...Total *
  teacher_student_usage_subset$Badges.per.Active.User
badges_per_teacher <- teacher_student_usage_subset %>% 
  group_by(Teacher.User.ID) %>% 
  summarize(Total.Badges = sum(Total.Badges))

# Statistics for "Total minutes on Zearn"
minutes_per_teacher <- teacher_usage %>%
  filter(!Adult.User.ID %in% unique_teachers) %>%
  group_by(Adult.User.ID) %>%
  summarize(Total.Minutes = sum(Minutes.on.Zearn...Total))

# Creating a function for common statistics
calculate_stats <- function(data, value) {
  list(
    N = sum(data[[value]]),
    Mean = mean(data[[value]]),
    SD = sd(data[[value]]),
    Quantiles = quantile(data[[value]], probs = c(0.25, 0.5, 0.75))
  )
}

# Calculating statistics
stats_classes <- calculate_stats(classes_per_teacher, "n")
stats_badges <- calculate_stats(badges_per_teacher, "Total.Badges")
stats_minutes <- calculate_stats(minutes_per_teacher, "Total.Minutes")

# Formatting the table with gt
gt_table <- tibble(
  Category = c("# of teachers", "# of classes", "# of badges", "Total minutes on Zearn"),
  N = c(n_teachers, stats_classes$N, stats_badges$N, stats_minutes$N),
  Mean = c(NA, stats_classes$Mean, stats_badges$Mean, stats_minutes$Mean),
  `Standard Deviation` = c(NA, stats_classes$SD, stats_badges$SD, stats_minutes$SD),
  `1st Quartile` = c(NA, stats_classes$Quantiles[1], stats_badges$Quantiles[1], stats_minutes$Quantiles[1]),
  `2nd Quartile` = c(NA, stats_classes$Quantiles[2], stats_badges$Quantiles[2], stats_minutes$Quantiles[2]),
  `3rd Quartile` = c(NA, stats_classes$Quantiles[3], stats_badges$Quantiles[3], stats_minutes$Quantiles[3])
) %>%
  gt() %>%
  tab_spanner(
    label = "Aggregated Statistics per Teacher",
    columns = c("Mean", "Standard Deviation", "1st Quartile", "2nd Quartile", "3rd Quartile")
  ) %>%
  cols_label(
    Category = "",
    N = "N",
    Mean = "Mean",
    `Standard Deviation` = "Standard Deviation",
    `1st Quartile` = "1st Quartile",
    `2nd Quartile` = "2nd Quartile",
    `3rd Quartile` = "3rd Quartile"
  ) %>%
  fmt_number(
    columns = c("N"),
    decimals = 0
  ) %>%
  fmt_number(
    columns = c("Mean", "Standard Deviation"),
    decimals = 2,
    use_seps = TRUE
  ) %>%
  fmt_number(
    columns = c("1st Quartile", "2nd Quartile", "3rd Quartile"),
    decimals = 1
  ) %>%
  sub_missing(
    columns = c("Mean", "Standard Deviation", "1st Quartile", "2nd Quartile", "3rd Quartile"),
    missing_text = "-"
  )

# Display the table
gt_table

```

### Study 1a: Data-driven Nudge Engineering with PCA

#### PCA Execution:
  -   Selection criteria for variables included in PCA.
  -   Step-by-step PCA process.
  -   Software and version used for PCA.

```{r}

# Perform PCA to choose number of components
# Removing columns with zero variance
ICA_cols <- sapply(teacher_student_usage_subset[13:39], function(x) sd(x) != 0)
pca_result <- prcomp(teacher_student_usage_subset[13:39][, ICA_cols],
                     scale. = TRUE)
prop_variance <- summary(pca_result)$importance[2, 1:10]
# Create Elbow Plot
elbow_plot <- ggplot(data.frame(Cluster = 1:10, Variance = prop_variance),
                     aes(x = Cluster, y = Variance)) +
  geom_line(color = "gray") +
  geom_point() +
  labs(x = "N-th Cluster", y = "Proportion of Variance",
       title = "Elbow Method for Determining Optimal k") +
  theme_minimal()

# Perform ICA
ica_result <- fastICA(X = teacher_student_usage_subset[13:39][, ICA_cols],
            n.comp = 3, row.norm = TRUE)
# Add ICA components to the dataset
teacher_student_usage_subset <- teacher_student_usage_subset %>%
  mutate(
    ica1 = ica_result$S[, 1],
    ica2 = ica_result$S[, 2],
    ica3 = ica_result$S[, 3]
  )

# Return ICA mixing matrix
ica_result$A

sapply(as.data.frame(as.matrix(ica_result$X)%*%t(as.matrix(ica_result$A))),
       FUN = var) /
  sum(sapply(as.data.frame(as.matrix(ica_result$X)%*%t(as.matrix(ica_result$A))),
             FUN = var))


## Robustness
# Exclude User.Session
ica_result2 <- fastICA(X = teacher_student_usage_subset[14:39][, ICA_cols[-1]],
            n.comp = 3, row.norm = TRUE)
# Return ICA mixing matrix
ica_result2$A

sapply(as.data.frame(as.matrix(ica_result2$X)%*%t(as.matrix(ica_result2$A))),
       FUN = var) /
  sum(sapply(as.data.frame(as.matrix(ica_result2$X)%*%t(as.matrix(ica_result2$A))),
             FUN = var))

```

#### **Statistical Considerations**:
  -   Explanation of statistical models used.
  -   Justification of model choices.
  -   Details of any sensitivity or robustness checks.
  -   Training on Statistical Methods: Detail any specific statistical training or standardized methodologies used.

### Study 1b: Data-driven Nudge Engineering with Habit-based Regression

#### **Data Utilization**:
    -   Criteria for data selection.
    -   Description of data preprocessing.
    -   Pre-registration of Study: If applicable, mention if this study was pre-registered.
#### **Regression Analysis Methodology**:
    -   Detailed steps in the regression analysis.
    -   Explanation of fixed effects model.
    -   Software and tools used in analysis.
    -   Publishing Negative Data: Include any negative data that did not support hypotheses.
#### **Statistical Tools and Software**:
    -   List of all statistical software and tools.
    -   Versions and settings used.
    -   Thorough Description of Methods: Clearly report experimental parameters.

### Study 2a: Nudge Engineering / Increasing Engagement Through Empathy

#### **Pre-registration Details**:
    -   Date created: September 14, 2021.
    -   Confirmation that no data were collected prior to the study.
    -   Hypotheses: H1 and H2 with detailed descriptions.
    -   Dependent Variables: Primary and secondary variables with specific definitions and timeframes.
    -   Conditions: Description of "Empathy" treatment and "Control Condition".
    -   Analysis Plan: Description of regression analysis, control variables, handling of missing values, and adjustments for multiple comparisons.
    -   Outliers and Exclusions: Specific criteria for exclusion of observations.
    -   Sample Size: Number of participants and criteria for eligibility.
    -   Additional Analyses: Detailed list of supplementary analyses to be conducted.
#### **Intervention Design**:
    -   Steps in developing empathy-based interventions.
    -   Criteria for intervention selection.
#### **Implementation Method**:
    -   Detailed procedure for implementing interventions.
    -   Timeline and phases of implementation.
#### **Evaluation Technique**:
    -   Methods of measuring intervention impact.
    -   Statistical analysis of results.
    -   Detailed Reporting of Methodologies: Include comprehensive description of all methodologies used.
#### **Outcome Measure**:
    -   Specific measures used to evaluate the impact of interventions.
#### **IRB Information**:
    -   Details of Institutional Review Board (IRB) approvals and ethical considerations, if applicable.

### Study 2b: Nudge Engineering / Increasing Engagement Through Habitization

#### **Pre-registration Details**:
    -   Date created: September 14, 2021.
    -   Confirmation that no data were collected prior to the study.
    -   Hypotheses: H1, H2, and H3 with detailed descriptions.
    -   Dependent Variables: Primary and secondary variables with specific definitions and timeframes.
    -   Conditions: Description of "Friday log-in", "Wednesday log-in", and "Control Condition".
    -   Analysis Plan: Description of regression analysis, control variables, handling of missing values, and adjustments for multiple comparisons.
    -   Outliers and Exclusions: Specific criteria for exclusion of observations.
    -   Sample Size: Number of participants and criteria for eligibility.
    -   Additional Analyses: Detailed list of supplementary analyses to be conducted.
#### **Nudge Development Process**:
    -   Detailed steps in nudge creation.
    -   Basis for selecting specific nudges.
#### **Application Protocol**:
    -   Procedure for nudge application.
    -   Monitoring and adaptation strategies.
#### **Impact Assessment Method**:
    -   Techniques for evaluating nudge effectiveness.
    -   Statistical methods for analyzing outcomes.
    -   Comprehensive Methodological Reporting: Ensure thorough reporting of all methods and experimental parameters.
#### **Outcome Measure**:
    -   Specific measures used to evaluate the impact of interventions.
#### **IRB Information**:
    -   Details of Institutional Review Board (IRB) approvals and ethical considerations, if applicable.

